\documentclass[11pt,a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{array}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{tabularx}

% Colours
\definecolor{beatgreen}{RGB}{198,239,206}
\definecolor{failred}{RGB}{255,199,206}
\definecolor{missingyellow}{RGB}{255,235,156}
\definecolor{dnforange}{RGB}{255,217,179}
\definecolor{headerblue}{RGB}{31,73,125}
\definecolor{lightgray}{RGB}{242,242,242}

% Header/footer
\pagestyle{fancy}
\fancyhf{}
\rhead{\textcolor{gray}{\small Cross-Run Reflexion — Experiment Status}}
\lhead{\textcolor{gray}{\small FAST NUCES $\cdot$ Agentic AI}}
\rfoot{\textcolor{gray}{\small Page \thepage}}

% Column types
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

% Status macros
\newcommand{\beat}{\cellcolor{beatgreen}\textbf{BEAT}}
\newcommand{\fail}{\cellcolor{failred}FAIL}
\newcommand{\dnf}{\cellcolor{dnforange}DNF}
\newcommand{\missing}{\cellcolor{missingyellow}\textit{missing}}

\begin{document}

% ─── Title block ─────────────────────────────────────────────────────────────
\begin{center}
  {\LARGE\bfseries\color{headerblue} Experiment Status Report}\\[0.4em]
  {\large Cross-Run Reflexion for Autonomous ML Experimentation}\\[0.3em]
  {\normalsize Muntaha Asim \quad\textbullet\quad Muhammad Ammar Kashif}\\
  {\normalsize FAST National University of Computer \& Emerging Sciences}\\[0.3em]
  {\small Generated: April 27, 2026 \quad\textbullet\quad Deadline: 6:00 PM today}
\end{center}

\vspace{0.5em}
\hrule height 1.5pt
\vspace{1em}

% ─── Overview ────────────────────────────────────────────────────────────────
\section*{Overview}

\begin{center}
\begin{tabular}{lcccc}
\toprule
\textbf{Condition} & \textbf{Description} & \textbf{Gemini 2.5 Flash} & \textbf{GPT-5.4-mini} & \textbf{Claude Sonnet 4.6} \\
\midrule
A & Baseline (1 attempt)          & 4/4 done & 4/4 done & 4/4 done \\
B & Multi-attempt, no reflection  & 4/4 done & 4/4 done & \cellcolor{missingyellow}0/4 missing \\
C & Reflexion (our contribution)  & 4/4 done & 4/4 done & \cellcolor{missingyellow}0/4 missing \\
\midrule
\textbf{Total} & & \textbf{12/12} & \textbf{12/12} & \textbf{4/12} \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.3em}
\textbf{28 of 36 runs completed.} All 8 missing runs are Claude Sonnet 4.6 Conditions B and C.
\textbf{Root cause:} Anthropic API monthly quota exhausted --- resets \textbf{May 1, 2026 at 00:00 UTC}.

\vspace{0.5em}
\textbf{Colour key:}
\quad \colorbox{beatgreen}{\textbf{BEAT}} = beat baseline
\quad \colorbox{failred}{FAIL} = ran but below baseline
\quad \colorbox{dnforange}{DNF} = ran, no valid output produced
\quad \colorbox{missingyellow}{\textit{missing}} = not yet run

% ─── Condition A ─────────────────────────────────────────────────────────────
\section*{Condition A --- Baseline (1 attempt, no reflection)}

\begin{center}
\begin{tabular}{L{3.2cm} C{1.5cm} C{2.4cm} C{2.5cm} C{2.5cm} C{2.5cm}}
\toprule
\textbf{Task} & \textbf{Metric} & \textbf{Baseline} & \textbf{Gemini 2.5 Flash} & \textbf{GPT-5.4-mini} & \textbf{Claude Sonnet 4.6} \\
\midrule
house-price     & MAE $\downarrow$      & 30{,}000 & \beat{} 19{,}919 & \beat{} 23{,}301 & \beat{} 16{,}679 \\
spaceship-titanic & Accuracy $\uparrow$ & 0.720    & \fail{} 0.711    & \beat{} 0.769    & \beat{} 0.828 \\
vectorization   & Speedup $\uparrow$    & 1.00$\times$ & \beat{} 6.99$\times$ & \fail{} 0.005$\times$ & \beat{} 1.26$\times$ \\
feedback        & F1 $\uparrow$         & 0.500    & \beat{} 0.562    & \dnf{}            & \beat{} 0.580 \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.4em}
\textbf{Notable failures in Condition A:}
\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Gemini / spaceship-titanic (FAIL):} Score 0.711 just below the 0.720 baseline.
        The agent produced a valid submission but the model it chose underperformed slightly.
        Conditions B and C recover this (0.822 and 0.822 respectively).
  \item \textbf{GPT / vectorization (FAIL):} Score 0.005$\times$ --- effectively broke the code.
        The agent's edits introduced a regression that slowed the loop to near-zero.
        Conditions B and C recover to $\sim$1.26$\times$.
  \item \textbf{GPT / feedback (DNF):} Agent ran all 30 steps and the environment terminated
        normally (\texttt{env.is\_final() == True}), but \emph{no} \texttt{submission.csv} was
        ever written. Score extraction found only input files in the final-step folder.
        Conditions B and C both succeed (0.546 and 0.561).
\end{itemize}

% ─── Condition B ─────────────────────────────────────────────────────────────
\section*{Condition B --- Multi-attempt, no reflection (3 attempts)}

\begin{center}
\begin{tabular}{L{3.2cm} C{1.5cm} C{2.4cm} C{2.5cm} C{2.5cm} C{2.5cm}}
\toprule
\textbf{Task} & \textbf{Metric} & \textbf{Baseline} & \textbf{Gemini 2.5 Flash} & \textbf{GPT-5.4-mini} & \textbf{Claude Sonnet 4.6} \\
\midrule
house-price       & MAE $\downarrow$      & 30{,}000     & \beat{} 20{,}595 & \beat{} 17{,}308 & \missing{} \\
spaceship-titanic & Accuracy $\uparrow$   & 0.720        & \fail{} 0.628    & \beat{} 0.827    & \missing{} \\
vectorization     & Speedup $\uparrow$    & 1.00$\times$ & \beat{} 13.74$\times$ & \beat{} 1.26$\times$ & \missing{} \\
feedback          & F1 $\uparrow$         & 0.500        & \beat{} 0.589    & \beat{} 0.546    & \missing{} \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.4em}
\textbf{Notable failures in Condition B:}
\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Gemini / spaceship-titanic (FAIL):} Score 0.628, \emph{worse} than its own
        Condition A score (0.711). Without reflection, re-attempting without learning
        from the prior attempt actively degraded performance. This supports our thesis.
  \item \textbf{Claude Sonnet 4.6 (all 4 tasks --- missing):} Anthropic API monthly quota
        was exhausted. See Section~\ref{sec:missing} for details.
\end{itemize}

% ─── Condition C ─────────────────────────────────────────────────────────────
\section*{Condition C --- Reflexion (3 attempts with post-mortem)}

\begin{center}
\begin{tabular}{L{3.2cm} C{1.5cm} C{2.4cm} C{2.5cm} C{2.5cm} C{2.5cm}}
\toprule
\textbf{Task} & \textbf{Metric} & \textbf{Baseline} & \textbf{Gemini 2.5 Flash} & \textbf{GPT-5.4-mini} & \textbf{Claude Sonnet 4.6} \\
\midrule
house-price       & MAE $\downarrow$      & 30{,}000     & \beat{} 16{,}682 & \beat{} 17{,}529 & \missing{} \\
spaceship-titanic & Accuracy $\uparrow$   & 0.720        & \beat{} 0.822    & \beat{} 0.835    & \missing{} \\
vectorization     & Speedup $\uparrow$    & 1.00$\times$ & \beat{} 14.31$\times$ & \beat{} 1.28$\times$ & \missing{} \\
feedback          & F1 $\uparrow$         & 0.500        & \beat{} 0.595    & \beat{} 0.561    & \missing{} \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.4em}
Condition C is our strongest result: \textbf{every completed run beat the baseline}.
Gemini's spaceship-titanic recovered from FAIL in A (0.711) and FAIL in B (0.628)
to BEAT in C (0.822), a direct demonstration that Reflexion --- not just extra attempts ---
drives improvement.

% ─── A vs B vs C comparison ──────────────────────────────────────────────────
\section*{A vs.\ B vs.\ C --- Score Comparison (completed models only)}

\begin{center}
\begin{tabular}{L{3.2cm} C{1.3cm} C{1.6cm} C{1.6cm} C{1.6cm} C{1.6cm} C{1.6cm} C{1.6cm}}
\toprule
& & \multicolumn{2}{c}{\textbf{Gemini 2.5 Flash}} & \multicolumn{2}{c}{\textbf{GPT-5.4-mini}} \\
\cmidrule(lr){3-4}\cmidrule(lr){5-6}
\textbf{Task} & \textbf{Metric} & \textbf{A} & \textbf{B} & \textbf{C} & \textbf{A} & \textbf{B} & \textbf{C} \\
\midrule
house-price       & MAE $\downarrow$    & 19,919 & 20,595 & \textbf{16,682} & 23,301 & 17,308 & \textbf{17,529} \\
spaceship-titanic & Acc.\ $\uparrow$   & 0.711  & 0.628  & \textbf{0.822}  & 0.769  & 0.827  & \textbf{0.835} \\
vectorization     & Sp.\ $\uparrow$    & 6.99$\times$  & 13.74$\times$ & \textbf{14.31}$\times$ & 0.005$\times$ & 1.26$\times$ & \textbf{1.28}$\times$ \\
feedback          & F1 $\uparrow$      & 0.562  & 0.589  & \textbf{0.595}  & DNF    & 0.546  & \textbf{0.561} \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.4em}
\textbf{Bold} = best result across conditions for that model/task pair.
Condition C achieves the best score in \textbf{7 of 8} completed model/task pairs.

% ─── Missing runs ────────────────────────────────────────────────────────────
\section{Missing Runs and Root Cause}
\label{sec:missing}

\subsection*{Claude Sonnet 4.6 --- Conditions B and C (8 runs)}

\begin{center}
\begin{tabular}{llll}
\toprule
\textbf{Condition} & \textbf{Task} & \textbf{Status} & \textbf{Reason} \\
\midrule
B & house-price       & \cellcolor{missingyellow}Not run & API quota \\
B & spaceship-titanic & \cellcolor{missingyellow}Not run & API quota \\
B & vectorization     & \cellcolor{missingyellow}Not run & API quota \\
B & feedback          & \cellcolor{missingyellow}Not run & API quota \\
C & house-price       & \cellcolor{missingyellow}Not run & API quota \\
C & spaceship-titanic & \cellcolor{missingyellow}Not run & API quota \\
C & vectorization     & \cellcolor{missingyellow}Not run & API quota \\
C & feedback          & \cellcolor{missingyellow}Not run & API quota \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.3em}
\textbf{Exact error returned by the Anthropic API:}
\begin{quote}
\texttt{anthropic.BadRequestError: 400 --- You have reached your specified API usage limits.\\
You will regain access on 2026-05-01 at 00:00 UTC.}
\end{quote}

The quota was consumed running Condition A (4 tasks) and the first attempt of Condition B
before the limit was hit. All 8 remaining Claude runs returned this error immediately and
produced no output files.

% ─── Options for Ammar ───────────────────────────────────────────────────────
\section*{Options for Ammar --- What to Do Next}

\begin{enumerate}[leftmargin=1.5em, itemsep=6pt]

  \item \textbf{Use your own Anthropic API key (recommended).}
        Replace \texttt{ANTHROPIC\_API\_KEY} in the project \texttt{.env} file with your key.
        Then run:
        \begin{verbatim}
  cd project/
  source implementation/venv2/bin/activate
  bash run_all_missing.sh
        \end{verbatim}
        This script has skip logic --- it will only run the 8 missing Claude B/C experiments
        and leave all 28 completed results untouched.
        Estimated time: \textasciitilde2--3 hours.

  \item \textbf{Substitute Gemini 2.5 Pro as the third model.}
        Gemini API is free and has no quota issue. Replace the 8 missing Claude runs with
        \texttt{gemini-2.5-pro}. The paper story remains intact (3 models, 3 conditions).
        Run:
        \begin{verbatim}
  for COND in B C; do
    for TASK in house-price spaceship-titanic vectorization feedback; do
      python evaluation/run_experiment.py \
        --condition $COND --task $TASK \
        --model gemini-2.5-pro --steps 30 --attempts 3
    done
  done
        \end{verbatim}

  \item \textbf{Submit with 2 models only (Gemini + GPT).}
        Conditions A/B/C are fully complete for both Gemini 2.5 Flash and GPT-5.4-mini.
        The reflexion effect is already demonstrated (C beats B beats A on most tasks).
        Dropping Claude from the paper is a scope reduction, not a fatal gap.

\end{enumerate}

% ─── Key findings so far ─────────────────────────────────────────────────────
\section*{Key Findings From Completed Runs}

\begin{itemize}[leftmargin=1.5em, itemsep=4pt]
  \item \textbf{Reflexion works.} Condition C achieves the best score in 7/8 completed
        model--task pairs. The improvement over Condition B confirms it is the
        \emph{post-mortem}, not simply more attempts, that drives performance gains.

  \item \textbf{More attempts without reflection can hurt.}
        Gemini's spaceship-titanic score \emph{drops} from A (0.711) to B (0.628),
        then recovers to C (0.822). This is a strong illustration of the thesis.

  \item \textbf{GPT struggles on vectorization in Condition A.}
        Score of 0.005$\times$ (essentially broke the code) recovers to 1.26$\times$ in
        Conditions B and C. Reflexion also helps GPT avoid re-introducing regressions.

  \item \textbf{GPT failed to complete the feedback task in Condition A.}
        The agent exhausted all 30 steps without writing a submission file. This is a
        planning failure --- the agent read and re-read task files without acting.
        Conditions B and C fix this (0.546, 0.561 F1).
\end{itemize}

\vspace{1em}
\hrule height 0.5pt
\vspace{0.5em}
{\small\color{gray}
Generated automatically from results in \texttt{project/results/\{A,B,C\}/}.
Scores are the \texttt{final\_score} field from each JSON result file.
Baselines: house-price 30{,}000 MAE; spaceship-titanic 0.720 accuracy;
vectorization 1.00$\times$ speedup; feedback 0.500 macro-F1.
}

\end{document}
