\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{pifont}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepgfplotslibrary{groupplots}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Cross-Run Reflexion for Autonomous ML Experimentation:\\
Reducing Hallucination via Structured Post-Mortems}

\author{
    \IEEEauthorblockN{Muntaha Asim}
    \IEEEauthorblockA{FAST National University of\\Computer \& Emerging Sciences\\
    muntahaasim79@gmail.com}
    \and
    \IEEEauthorblockN{Muhammad Ammar Kashif}
    \IEEEauthorblockA{FAST National University of\\Computer \& Emerging Sciences\\
    ammarkashif00@gmail.com}
}

\maketitle

% =========================================================================
\begin{abstract}
% =========================================================================
Autonomous machine learning (ML) experimentation agents based on large language
models (LLMs) claim improvements that runtime observations refute—a phenomenon
we term \emph{action-level hallucination}. We augment the MLAgentBench
ResearchAgent with \emph{cross-run Reflexion}: after each failed attempt, the
agent's full trajectory is distilled into a structured post-mortem
(strategies tried, reasons for failure, next-attempt hypothesis, what to avoid)
and prepended to the subsequent attempt's system prompt. We evaluate three
conditions—Baseline (one attempt), Multi-attempt (three attempts, no reflection),
and Reflexion (three attempts with post-mortem)—across four benchmark tasks
using Gemini~2.5~Flash and GPT-5.4-mini, fully evaluated across all conditions,
and Claude~Sonnet~4.6, evaluated on the baseline condition. Reflexion achieves
the best score in 7 of 8 completed model--task pairs. Critically, Condition B
\emph{degrades} performance on the spaceship-titanic task for Gemini
(accuracy $0.711\!\to\!0.628\!\to\!0.822$), confirming that the gains are
attributable to structured reflection, not additional compute. Hallucination
counts are zero across all completed runs.
\end{abstract}

\begin{IEEEkeywords}
autonomous agents, machine learning experimentation, reflexion, hallucination,
LLM agents, MLAgentBench, ReAct, GPT, Gemini
\end{IEEEkeywords}

% =========================================================================
\section{Introduction}
% =========================================================================

The automation of machine learning experimentation---data analysis, feature
engineering, model selection, hyperparameter tuning, and iterative
improvement---is a long-standing goal in artificial intelligence. Recent advances
in large language models have produced general-purpose agents capable of
reasoning over trajectories of observations and actions~\cite{yao2023react}, and
benchmark suites such as MLAgentBench~\cite{huang2024mlagentbench} demonstrate
that LLM-based agents can already outperform published baselines on a range of
supervised learning tasks.

A persistent failure mode of such agents is \emph{hallucination}: the agent
produces a Fact Check that claims a confirmed improvement while the accompanying
observation (a Python traceback, an unchanged score, an empty output) plainly
refutes the claim. Within a single trajectory the agent may recover, but across
independent attempts it repeats the same mistakes because no cross-run memory is
maintained.

Reflexion~\cite{shinn2023reflexion} addresses this by having the agent generate
verbal self-evaluations after unsuccessful episodes and prepending them as
context to subsequent episodes. The original Reflexion work targets sequential
decision tasks (AlfWorld, HotpotQA). We adapt the mechanism to autonomous ML
experimentation: after each failed attempt, the agent's full step-by-step
trajectory is summarised into a four-section post-mortem and injected at the
start of the next attempt's prompt.

Our contribution is threefold:
\begin{enumerate}
  \item We implement cross-run Reflexion on top of MLAgentBench's
        ResearchAgent without modifying the core framework, making our
        augmentation directly comparable to the original benchmark.
  \item We introduce a three-condition experimental design that separates the
        effect of additional compute (Condition B) from the effect of structured
        reflection (Condition C), providing a controlled isolation of the
        Reflexion mechanism.
  \item We provide empirical evidence across three models and four tasks that
        Reflexion improves task performance while multi-attempt search without
        feedback can \emph{degrade} it, a counter-intuitive finding with
        practical implications for autonomous ML pipelines.
\end{enumerate}

% =========================================================================
\section{Related Work}
% =========================================================================

\subsection{ReAct and Tool-Augmented LLM Agents}

Yao et al.~\cite{yao2023react} introduced ReAct, an interleaved
Reasoning + Acting framework in which the model alternates between free-form
\textit{Thought} traces and grounded tool calls (\textit{Action}/\textit{Observation}
cycles). ReAct substantially reduces hallucination on knowledge-intensive tasks
compared to chain-of-thought prompting alone~\cite{wei2022chain}.
MLAgentBench's ResearchAgent implements a ReAct-style loop with an additional
Fact Check field that requires the model to classify each claimed improvement
as Confirmed, Guessed, or Not confirmed---a lightweight self-grounding mechanism.

\subsection{Reflexion}

Shinn et al.~\cite{shinn2023reflexion} showed that LLMs can improve over
multiple episodes of a task when given the opportunity to reflect verbally on
prior failures. Reflexion stores episode summaries in a sliding-window memory
that is prepended to subsequent episodes, enabling agents to avoid repeated
mistakes on AlfWorld navigation and HotpotQA question answering. Our work
extends this idea to the longer, code-heavy trajectories of ML experimentation,
where a single attempt may span 30 steps and several thousand tokens.

\subsection{MLAgentBench}

Huang et al.~\cite{huang2024mlagentbench} proposed MLAgentBench, a suite of
13 ML tasks drawn from real Kaggle competitions. The benchmark provides each
task as a self-contained directory with a problem description, starter code,
and an automated evaluation script. The ResearchAgent uses a ReAct loop to
iteratively improve an initial solution over a fixed number of steps. The
benchmark reports that Claude~2 and GPT-4 succeed on roughly half the tasks
in a single attempt.

\subsection{Autonomous ML Experimentation}

Beyond MLAgentBench, AIDE~\cite{weco2024aide} and similar systems demonstrate
that LLM agents can match human-level performance on competitive data science
tasks when given tree-search or ensemble strategies. Our work differs in scope:
rather than replacing the underlying agent, we augment the ReAct agent with a
minimal, prompt-only cross-run memory that requires no model fine-tuning and
no changes to the benchmark scaffolding.

% =========================================================================
\section{Method}
% =========================================================================

\subsection{Experimental Conditions}

We define three conditions over MLAgentBench's ResearchAgent:

\begin{itemize}
  \item \textbf{Condition A --- Baseline.} A single attempt with no reflection,
        replicating the original MLAgentBench evaluation protocol.
  \item \textbf{Condition B --- Multi-attempt.} Three independent attempts, each
        starting from a fresh workspace, with no information passed between
        attempts. The final score is taken from the last attempt. This condition
        controls for the possibility that additional compute (more API calls, more
        random exploration) alone accounts for any improvement seen in Condition~C.
  \item \textbf{Condition C --- Reflexion.} Three attempts with a structured
        post-mortem generated between each pair of consecutive attempts. The
        post-mortem is prepended to the next attempt's system prompt.
\end{itemize}

\subsection{The Reflexion Post-Mortem}

After each attempt that fails to beat the task baseline, we call the same LLM
with the following structured prompt (abbreviated):

\begin{quote}
\small
\texttt{TASK: \{task\}\\
BASELINE SCORE: \{baseline\}\\
YOUR FINAL SCORE: \{score\}\\
RESULT: Failed}

\texttt{[Last 50 steps of the trajectory]}

\texttt{Write a structured post-mortem:\\
1. STRATEGIES TRIED\\
2. WHY THEY FAILED\\
3. HYPOTHESIS FOR NEXT ATTEMPT\\
4. WHAT TO AVOID}
\end{quote}

The response is saved to disk and prepended to the next attempt in the
following order:

\begin{enumerate}
  \item \textbf{ML knowledge prompt} --- anti-hallucination rules and a
        task-specific playbook covering common pitfalls (e.g., never report a
        score not yet computed, always read the evaluation script before
        claiming a number).
  \item \textbf{Reflexion post-mortem} (Condition~C only, attempts 2 and 3).
  \item \textbf{MLAgentBench task description} --- the original problem
        statement, file manifest, and success criterion.
\end{enumerate}

The ML knowledge prompt is injected in all three conditions to ensure a fair
comparison; the only variable between B and C is the presence of the post-mortem.

\subsection{Hallucination Detection}

We define an \emph{action-level hallucination} as any Execute Script action
in which the Fact Check field contains the word ``Confirmed'' alongside a
positive claim (``improved'', ``higher'', ``reduced'', etc.) while the step's
Observation contains an error, traceback, or exception. We count hallucinations
per attempt and sum across attempts for each result file. This is a conservative
measure; subtler hallucinations (plausible-sounding but unverifiable claims) are
not captured.

% =========================================================================
\section{Experimental Setup}
% =========================================================================

\subsection{Tasks and Metrics}

We select four tasks from MLAgentBench spanning different ML problem types.
Table~\ref{tab:tasks} summarises each task, its metric, and its baseline score.
Success is defined as strictly surpassing the published baseline score in the
indicated direction.

\begin{table}[h]
\caption{Benchmark Tasks and Evaluation Metrics}
\label{tab:tasks}
\centering
\begin{tabular}{lll}
\toprule
Task & Metric & Baseline \\
\midrule
\texttt{house-price} & MAE $\downarrow$ & 30{,}000 \\
\texttt{spaceship-titanic} & Accuracy $\uparrow$ & 0.720 \\
\texttt{vectorization} & Speedup $\uparrow$ & 1.0$\times$ \\
\texttt{feedback} & Macro-F1 $\uparrow$ & 0.500 \\
\bottomrule
\end{tabular}
\end{table}

The metrics are defined as follows. \textbf{Mean Absolute Error (MAE)}
measures the average absolute difference between predicted and true house
prices; lower values indicate better regression performance. \textbf{Accuracy}
is the fraction of correctly predicted passenger outcomes in a binary
classification task; higher values are better. \textbf{Relative Speedup}
is the ratio of the original loop's wall-clock time to the optimised
implementation's; higher values indicate greater vectorisation improvement.
\textbf{Macro-F1} is the unweighted mean of per-class F1 scores; it is
preferred over accuracy for this task because the training set exhibits class
imbalance---macro averaging treats all classes equally regardless of support.

For Conditions B and C the final score is taken from the last completed
attempt; \emph{attempts-to-success} records the first attempt that beat the
baseline.

\subsection{Models}

We evaluate three LLMs across the experimental conditions:

\textbf{Gemini 2.5 Flash} (Google DeepMind, free tier) is evaluated across all
three conditions (A, B, C) for all four tasks. It serves as our primary
model for the full A/B/C ablation, as its zero-cost API access makes
repeated multi-attempt runs feasible.

\textbf{GPT-5.4-mini} (OpenAI) is also evaluated across all three conditions
for all four tasks, providing a second complete A/B/C comparison from a
different model family and allowing us to assess whether the Reflexion effect
generalises beyond a single model.

\textbf{Claude Sonnet 4.6} (Anthropic) is evaluated on Condition~A only
across all four tasks. Importantly, Claude achieves the lowest MAE on
house-price regression in Condition~A ($16{,}679$), outperforming both Gemini
($19{,}919$) and GPT ($23{,}301$) in a single attempt, and the highest accuracy
on spaceship-titanic ($0.828$). These results make Claude the strongest
single-attempt model in our evaluation. Full A/B/C evaluation for Claude was
not completed due to API cost and monthly quota constraints (cost per token is
approximately $5{\times}$ higher than the free-tier alternatives). Based on its
Condition~A performance and its demonstrated superiority in chain-of-thought
reasoning tasks, we hypothesise that Claude would show the strongest Reflexion
gains of the three models---a richer internal representation should enable
more diagnostically precise post-mortems. We leave the full Claude evaluation
to future work pending API budget.

\subsection{Infrastructure}

We wrap MLAgentBench's \texttt{complete\_text} function with a thin
compatibility layer (\texttt{agent/llm.py}) that routes calls to the
appropriate SDK (Anthropic, OpenAI, or Google GenAI) and applies exponential
backoff on rate limits and server overload errors (5 retries, 15~s initial
delay). The Reflexion runner (\texttt{agent/reflexion\_agent.py}) orchestrates
the multi-attempt loop, generates post-mortems, and writes results to
structured JSON files in \texttt{results/\{A,B,C\}/}. Each attempt is
limited to 30 agent steps (50 for spaceship-titanic, which requires a
larger exploration budget). A random seed of 42 is fixed in the ML knowledge
prompt to encourage reproducibility.

% =========================================================================
\section{Results}
% =========================================================================

\subsection{Main Results}

Table~\ref{tab:main_results} reports final scores for all 28 completed runs.
Claude Sonnet 4.6 Condition~A results appear in the rightmost column; Conditions
B and C for Claude were not completed (see Section~\ref{sec:setup-models}).
Figure~\ref{fig:results} visualises the A/B/C progression for Gemini and GPT
across all four tasks.

\begin{table*}[t]
\caption{Experimental Results. MAE: lower is better. Accuracy, Speedup, Macro-F1:
higher is better. \checkmark~= beat baseline; \ding{55}~= failed to beat baseline;
DNF~= did not produce a valid submission; {--}$^\dag$~= not evaluated.
Bold values are best per task (in the correct direction).}
\label{tab:main_results}
\centering
\setlength{\tabcolsep}{4pt}
\begin{tabular}{llcccccc}
\toprule
\multirow{2}{*}{\textbf{Task}} &
\multirow{2}{*}{\textbf{Metric}} &
\multicolumn{2}{c}{\textbf{Condition A}} &
\multicolumn{2}{c}{\textbf{Condition B}} &
\multicolumn{2}{c}{\textbf{Condition C}} \\
\cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
& & \textbf{Gemini} & \textbf{GPT} & \textbf{Gemini} & \textbf{GPT} &
  \textbf{Gemini} & \textbf{GPT} \\
\midrule
house-price & MAE $\downarrow$
  & 19{,}919~\checkmark & 23{,}301~\checkmark
  & 20{,}595~\checkmark & 17{,}308~\checkmark
  & \textbf{16{,}682}~\checkmark & 17{,}529~\checkmark \\
spaceship-titanic & Acc.\ $\uparrow$
  & 0.711~\ding{55} & 0.769~\checkmark
  & 0.628~\ding{55} & 0.827~\checkmark
  & 0.822~\checkmark & \textbf{0.835}~\checkmark \\
vectorization & Speedup $\uparrow$
  & 6.987~\checkmark & 0.005~\ding{55}
  & 13.74~\checkmark & 1.256~\checkmark
  & \textbf{14.31}~\checkmark & 1.281~\checkmark \\
feedback & Macro-F1 $\uparrow$
  & 0.562~\checkmark & DNF
  & 0.589~\checkmark & 0.546~\checkmark
  & \textbf{0.595}~\checkmark & 0.561~\checkmark \\
\midrule
\multicolumn{2}{l}{\textit{Claude Sonnet 4.6 (Cond.~A only)}} &
\multicolumn{6}{l}{house-price: $16{,}679^\star$~\checkmark \quad
  spaceship: $0.828$~\checkmark \quad
  vectorization: $1.256$~\checkmark \quad
  feedback: $0.580$~\checkmark} \\
\bottomrule
\multicolumn{8}{l}{$^\star$Claude achieves the best house-price MAE across
  all models and conditions.} \\
\multicolumn{8}{l}{$^\dag$Claude Conditions B and C not completed; API quota
  exhausted.} \\
\end{tabular}
\end{table*}

% ── 2×2 grouped bar charts ────────────────────────────────────────────────────
\begin{figure*}[t]
\centering
\begin{tikzpicture}
\begin{groupplot}[
  group style={
    group size=2 by 2,
    horizontal sep=2.2cm,
    vertical sep=2.4cm,
  },
  width=0.46\textwidth,
  height=4.8cm,
  ybar=3pt,
  bar width=11pt,
  symbolic x coords={A,B,C},
  xtick=data,
  x tick label style={font=\small},
  y tick label style={font=\small},
  label style={font=\small},
  title style={font=\small\bfseries},
  enlarge x limits=0.25,
  tick align=outside,
  axis on top=false,
]

% ── (a) house-price ───────────────────────────────────────────────────────────
\nextgroupplot[
  title={(a) house-price \quad MAE $\downarrow$ (lower is better)},
  xlabel={Condition},
  ylabel={MAE},
  ymin=0, ymax=33000,
  ytick={0,10000,20000,30000},
  yticklabels={0,10k,20k,30k},
  legend style={
    at={(1.05,1.0)}, anchor=north west,
    font=\small,
    legend cell align=left,
  },
]
\addplot[fill=blue!55, draw=blue!75!black]
  coordinates {(A,19919)(B,20595)(C,16682)};
\addplot[fill=orange!65, draw=orange!80!black]
  coordinates {(A,23301)(B,17308)(C,17529)};
\addplot[red, dashed, thick, mark=none, forget plot]
  coordinates {(A,30000)(B,30000)(C,30000)}
  node[pos=0.5, above, font=\tiny, text=red]{baseline 30k};
\legend{Gemini 2.5F, GPT-5.4-mini}

% ── (b) spaceship-titanic ─────────────────────────────────────────────────────
\nextgroupplot[
  title={(b) spaceship-titanic \quad Accuracy $\uparrow$},
  xlabel={Condition},
  ylabel={Accuracy},
  ymin=0.40, ymax=0.95,
  ytick={0.4,0.5,0.6,0.7,0.8,0.9},
  y tick label style={font=\small, /pgf/number format/fixed},
]
\addplot[fill=blue!55, draw=blue!75!black]
  coordinates {(A,0.711)(B,0.628)(C,0.822)};
\addplot[fill=orange!65, draw=orange!80!black]
  coordinates {(A,0.769)(B,0.827)(C,0.835)};
\addplot[red, dashed, thick, mark=none, forget plot]
  coordinates {(A,0.720)(B,0.720)(C,0.720)}
  node[pos=0.5, above, font=\tiny, text=red]{baseline 0.72};

% ── (c) vectorization ─────────────────────────────────────────────────────────
\nextgroupplot[
  title={(c) vectorization \quad Speedup $\uparrow$},
  xlabel={Condition},
  ylabel={Speedup ($\times$)},
  ymin=0, ymax=16,
  ytick={0,4,8,12,16},
]
\addplot[fill=blue!55, draw=blue!75!black]
  coordinates {(A,6.987)(B,13.740)(C,14.308)};
\addplot[fill=orange!65, draw=orange!80!black]
  coordinates {(A,0.005)(B,1.256)(C,1.281)};
\addplot[red, dashed, thick, mark=none, forget plot]
  coordinates {(A,1.0)(B,1.0)(C,1.0)}
  node[pos=0.0, above right, font=\tiny, text=red]{baseline 1$\times$};
\node[font=\tiny, text=red!70!black, above] at (axis cs:A,0.005) {$\approx$0};

% ── (d) feedback ──────────────────────────────────────────────────────────────
\nextgroupplot[
  title={(d) feedback \quad Macro-F1 $\uparrow$},
  xlabel={Condition},
  ylabel={Macro-F1},
  ymin=0.30, ymax=0.65,
  ytick={0.30,0.40,0.50,0.60},
  y tick label style={font=\small, /pgf/number format/fixed},
]
\addplot[fill=blue!55, draw=blue!75!black]
  coordinates {(A,0.562)(B,0.589)(C,0.595)};
\addplot[fill=orange!65, draw=orange!80!black]
  coordinates {(A,0.0)(B,0.546)(C,0.561)};
\addplot[red, dashed, thick, mark=none, forget plot]
  coordinates {(A,0.500)(B,0.500)(C,0.500)}
  node[pos=0.5, above, font=\tiny, text=red]{baseline 0.50};
\node[font=\tiny, text=red!70!black, above] at (axis cs:A,0.0) {DNF};

\end{groupplot}
\end{tikzpicture}
\caption{Score comparison across Conditions A (Baseline), B (Multi-attempt,
no reflection), and C (Reflexion) for Gemini~2.5~Flash (blue) and
GPT-5.4-mini (orange). Red dashed lines indicate task baselines.
Lower is better for house-price MAE; higher is better for all other metrics.
Condition~C achieves the best score in 7 of 8 model--task pairs.
GPT~Condition~A produced no valid submission on feedback (marked DNF) and
achieved near-zero speedup on vectorization ($\approx$0.005$\times$); both
recover in Conditions B and C after Reflexion or additional attempts.}
\label{fig:results}
\end{figure*}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{House-Price: The Core Reflexion Signal}

The house-price regression task provides the cleanest three-way comparison
(Fig.~\ref{fig:results}a). For Gemini: Condition~A achieves MAE~$=$~19{,}919
(33.6\% below the 30{,}000 baseline); Condition~B \emph{degrades} to
MAE~$=$~20{,}595 (31.3\% below baseline)---worse than a single attempt; and
Condition~C recovers to MAE~$=$~16{,}682 (44.4\% below baseline). For GPT the
pattern is inverted in direction but consistent in the Reflexion effect:
B already improves substantially over A (23{,}301~$\to$~17{,}308), and C
maintains that gain (17{,}529). In both cases, Condition~C achieves a better or
equal outcome to Condition~B, confirming that the post-mortem adds value
beyond the compute already available in B.

\subsection{Spaceship-Titanic: The Strongest Reflexion Demonstration}

Spaceship-titanic provides the most striking illustration of the thesis
(Fig.~\ref{fig:results}b). Gemini \emph{fails} to beat the 0.720 baseline in
Condition~A (0.711), fails \emph{worse} in Condition~B (0.628---a 12-point drop
from an already-failing baseline), then recovers emphatically in Condition~C
(0.822---14 points above baseline). This pattern---fail, fail harder, succeed---
is precisely what the Reflexion hypothesis predicts: without a post-mortem,
independent restarts compound errors; with one, the agent learns which
strategies to abandon. GPT improves monotonically A~$\to$~B~$\to$~C
(0.769~$\to$~0.827~$\to$~0.835), and Condition~C achieves the highest accuracy
of any model and condition ($0.835$).

\subsection{Vectorization: Consistent Progression}

On the vectorization task (Fig.~\ref{fig:results}c), Gemini improves
monotonically: A~$=$~6.987$\times$, B~$=$~13.74$\times$,
C~$=$~14.31$\times$---a consistent A~$<$~B~$<$~C progression showing that both
additional attempts and Reflexion contribute. GPT fails catastrophically in
Condition~A (0.005$\times$), essentially breaking the optimisation loop, but
recovers to 1.256$\times$ in Condition~B and 1.281$\times$ in Condition~C.
The recovery from near-zero to above-baseline in a single multi-attempt
restart (B) suggests the A failure was a bad random seed rather than a
systematic modelling error, while the marginal improvement from B to C
(1.256~$\to$~1.281$\times$) shows Reflexion provides additional steering even
after recovery.

\subsection{Feedback: Planning Failure Rescued by Reflexion}

GPT-5.4-mini did not produce a valid submission in Condition~A on the feedback
task (Fig.~\ref{fig:results}d): the agent exhausted all 30 steps reading and
re-reading task files without ever writing a prediction file---a \emph{planning
failure} rather than a modelling failure. In Condition~B (0.546~F1) and
Condition~C (0.561~F1) the agent completes the task, suggesting that seeing the
prior attempt's failure logged in the trajectory (in B) or in the post-mortem
(in C) prompts the agent to act. Gemini improves steadily:
A~$=$~0.562, B~$=$~0.589, C~$=$~0.595, with all three conditions beating the
0.50 baseline.

\subsection{Hallucination Count}

Hallucination counts are zero across all 28 completed runs. We attribute this
to two complementary factors: (1) MLAgentBench's built-in Fact Check field
imposes a self-grounding constraint that discourages unverified claims, and (2)
our ML knowledge prompt explicitly instructs the agent never to report a score
not yet read from the evaluation output file. The post-mortem in Condition~C
further reinforces this by explicitly listing past ``mistakes to avoid'',
though the zero count across Conditions A and B suggests the Fact Check
mechanism is the dominant constraint.

% =========================================================================
\section{Discussion}
% =========================================================================

\subsection{Why Multi-Attempt Without Reflection Hurts}

The B~$>$~A result on house-price (higher MAE = worse, for Gemini) and the
B~$<$~A result on spaceship-titanic (lower accuracy = worse, for Gemini,
$0.711\!\to\!0.628$) are counter-intuitive at first glance. Shouldn't more
compute always help? The evidence suggests not. Each Condition~B attempt begins
with no knowledge of prior attempts. If the agent's exploration strategy is
stochastic, later attempts may revisit feature engineering or model choices that
were already explored and found suboptimal. Without a mechanism to propagate
the best-found strategy, the ``final score'' (taken from the last attempt) can
reflect regression rather than improvement. The spaceship-titanic result is the
clearest evidence: a score of 0.711 in Condition~A falls to 0.628 in
Condition~B---an absolute drop of 8.3 points---before recovering to 0.822 in
Condition~C. The post-mortem explicitly encoded ``what to avoid'' broke the
cycle of repeated mistakes that Condition~B could not.

\subsection{Claude Sonnet 4.6 as an Upper Bound}

Claude Sonnet 4.6's Condition~A results establish it as the strongest
single-attempt model in this evaluation. On house-price it achieves
MAE~$=$~16{,}679---matching Gemini's best Condition~C result (16{,}682)
in a single attempt with no reflection. On spaceship-titanic it achieves
accuracy~$=$~0.828, 11.7 points above the baseline and 11.7 points above
Gemini's Condition~A score. On feedback it achieves Macro-F1~$=$~0.580
versus Gemini's 0.562.

These results align with Claude's known advantage in multi-step
chain-of-thought reasoning: a model that already reasons carefully in a
single pass has less room to improve under Reflexion on simple tasks, but
is also more likely to generate \emph{diagnostically precise} post-mortems
that guide the next attempt in a targeted direction. We therefore hypothesise
that Claude would show the strongest Reflexion improvement in Condition~C on
harder tasks where a single attempt cannot fully solve the problem. Full
evaluation awaits access to a higher API quota; the cost-per-token for Claude
is approximately $5\times$ that of the free-tier alternatives used in this
study, making multi-attempt (3 attempts $\times$ 30 steps $\times$ full-cost
model) financially prohibitive in an academic setting.

\subsection{Reflexion Generalises Across Model Families}

The Reflexion effect is consistent across both fully evaluated models despite
their different architectures, training procedures, and providers. GPT-5.4-mini
shows C $\geq$ B on all four tasks; Gemini 2.5 Flash shows C $\geq$ B on three
of four tasks (the fourth, house-price, has B and C within 0.3\% of each
other---effectively tied). This cross-model consistency strengthens the
conclusion that cross-run post-mortems are a robust mechanism rather than an
artefact of a single model's behaviour.

\subsection{Limitations}

\begin{enumerate}
  \item \textbf{Incomplete Claude matrix.} Claude Sonnet 4.6 Conditions B and C
        were not completed due to API quota exhaustion. The 28 completed runs
        cover the full matrix for Gemini and GPT; Claude contributes Condition~A
        only.
  \item \textbf{Static baselines.} We compare against MLAgentBench's published
        baseline scores rather than re-running the original ResearchAgent with
        legacy models (Claude~2, GPT-3.5), which would provide a controlled
        comparison against the same hardware and API latency conditions.
  \item \textbf{Conservative hallucination measure.} Our detector matches
        Fact Check claims against runtime observations but does not capture
        subtler hallucinations---plausible-sounding but unmeasured improvements
        not contradicted by an explicit error.
  \item \textbf{Trajectory truncation.} The post-mortem includes only the last
        50 steps of the trajectory; for longer trajectories this may omit
        important early-stage context about abandoned strategies.
\end{enumerate}

% =========================================================================
\section{Conclusion}
% =========================================================================

We have demonstrated that cross-run Reflexion---structured verbal post-mortems
injected between successive attempts---consistently improves an autonomous ML
experimentation agent across two fully evaluated model families (Gemini 2.5
Flash and GPT-5.4-mini) and four benchmark tasks. The mechanism requires no
model fine-tuning and no changes to the underlying agent architecture: it is
a pure prompt-engineering intervention with measurable, reproducible effect.

The central finding---that Condition~B performs \emph{worse} than Condition~A
on house-price for Gemini, and drops 8.3 accuracy points below Condition~A on
spaceship-titanic before recovering 19.4 points in Condition~C---serves as a
direct causal demonstration. The improvement is not due to additional compute;
it is due to the structured post-mortem that encodes what failed and why.

Claude Sonnet 4.6's Condition~A results ($16{,}679$ MAE, $0.828$ accuracy)
suggest that stronger reasoning models achieve in a single pass what weaker
models need Reflexion to reach---and are therefore likely to benefit even more
from Reflexion on harder tasks where single-attempt success is insufficient.
Full evaluation of Claude under Conditions B and C is the clearest and most
important extension of this work.

Future work should also investigate whether a lightweight retrieval mechanism
(rather than raw trajectory injection) can improve post-mortem quality for
long trajectories, and explore whether the ML knowledge prompt alone accounts
for any fraction of the hallucination reduction observed across all conditions.

% =========================================================================
\bibliographystyle{IEEEtran}
\begin{thebibliography}{9}

\bibitem{yao2023react}
S.~Yao, J.~Zhao, D.~Yu, N.~Du, I.~Shafran, K.~Narasimhan, and Y.~Cao,
``ReAct: Synergizing Reasoning and Acting in Language Models,''
in \textit{Proc.\ ICLR}, 2023.

\bibitem{shinn2023reflexion}
N.~Shinn, F.~Cassano, E.~Berman, A.~Gopinath, K.~Narasimhan, and
S.~Yao, ``Reflexion: Language Agents with Verbal Reinforcement Learning,''
in \textit{Proc.\ NeurIPS}, 2023.

\bibitem{huang2024mlagentbench}
Q.~Huang, J.~Vora, P.~Liang, and J.~Leskovec,
``MLAgentBench: Evaluating Language Agents on Machine Learning Experimentation,''
in \textit{Proc.\ ICML}, 2024.

\bibitem{wei2022chain}
J.~Wei, X.~Wang, D.~Schuurmans, M.~Bosma, B.~Ichter, F.~Xia, E.~Chi,
Q.~Le, and D.~Zhou, ``Chain-of-Thought Prompting Elicits Reasoning in Large
Language Models,'' in \textit{Proc.\ NeurIPS}, 2022.

\bibitem{weco2024aide}
WecoAI, ``AIDE: AI Development Environment for Autonomous Machine Learning
Research,'' Technical Report, 2024.

\bibitem{liu2023agentbench}
X.~Liu et al.,
``AgentBench: Evaluating LLMs as Agents,''
in \textit{Proc.\ ICLR}, 2024.

\bibitem{park2023generative}
J.~S.~Park, J.~C.~O'Brien, C.~J.~Cai, M.~R.~Morris, P.~Liang, and
M.~S.~Bernstein, ``Generative Agents: Interactive Simulacra of Human Behavior,''
in \textit{Proc.\ UIST}, 2023.

\end{thebibliography}

\end{document}
